---
title: "Stat4601_Bronx_Kmeans"
author: "Stephanie Cheng"
date: "2025-04-09"
output: pdf_document
---
```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(stats)
library(here)
library(ggplot2)

source("UL_helpers.r")

load("dataset_RData_cluster/bronx_data.RData")
```

# K-mean with PCA
Reduce dimensions and prepare data for clustering
```{r PCA}
bronx_pca <- pca(bronx_data, "Bronx")
```

Calculate clustering evaluation with Davies Bouldin index & Within-cluster sum of squares.
See the affect when K is increasing, then we can apply elbow method to avoid 
picking the best k within overfitting case. 
```{r Clustering Metrics}
bronx_k_stats_20 <- calculate_k_stats_PCA(bronx_pca, max_k = 20)
bronx_k_stats_40 <- calculate_k_stats_PCA(bronx_pca, max_k = 40)

# DBI & WSS plot
elbows_20 <- plot_kmeans(bronx_k_stats_20$errs, bronx_k_stats_20$DBI)
elbows_40 <- plot_kmeans(bronx_k_stats_40$errs, bronx_k_stats_40$DBI)
```

```{r Best K cluster}
best_k <- 5
```

Plot all clusters from 3 to 8 as the best k clusters is within that range. 
```{r K Clustering Plot}
plot_clusters(bronx_k_stats_20$X.syn, min_k = 3, max_k = 8)
```

K-means on PCA as PCA gives a lower-dimensional variable that improves
clustering quality
```{r perform k with Optimal K}
km <- kmeans(bronx_pca$x, centers = best_k, nstart = 25)
summarize_kmeans(km, "Bronx")
```

Interpret what the clusters mean with the original data
```{r Summarize Cluster}
bronx_data$cluster <- km$cluster
aggregate(. ~ cluster, data = bronx_data, mean)
```

Export the clusters for Supervised learning 
```{r export}
dir.create("after_cluster_dataset")
write.csv(bronx_data, 
          file = "after_cluster_dataset/cleaned_bronx_with_clusters.csv", 
          row.names = FALSE)
```
